Boosting Higgs Pair Production in the $bb\bar{b}\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-07-07 paper/sec-analysis.tex
work on paper

In section we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to the event topology.

First of all we discuss the settings and nomenclature that
we will use for jet clustering in this work.
%
Then we move to discuss how $b$-tagging is simulated,
following closely the ATLAS techniques.
%
Then we discuss the categorisation of events between different
topologies, and how we can prioritise among them.

@@

In our analysis we will use different types of jet definitions,
depending of the type of analysis that needs to be performed.
%
Therefore, we being by enumerating the choices that we make.
%
The starting point after the parton shower is the clustering
of all final state particles.
%
In this work all the jet reconstruction algorithms
are obtained from the {\tt FastJet} program~\cite{Cacciari:2011ma},
v3.1.0.
%
We adopt the following terminology for the different types of jets
that we will use in this work:

@@

  These are jets  reconstructed with the
  anti-$k_T$ clustering algorithm~\cite{Cacciari:2008gp} with $R=0.4$ radius.
  %
  When reconstructing small-$R$ jets, we will keep only
  those with transverse momentum $p_T \ge 40$~GeV
  and pseudo-rapidity $|\eta|<2.5$ within the central
  acceptance of ATLAS and CMS, jets that fail to satisfy these
  conditions are discarded

  They are required to have $p_T > 50$~GeV and $|\eta|<2.5$.

@@

  In addition to the basic $p_T$ and $\eta$
  selection requirements, large-$R$ jets should also
  satisfy the  BDRS mass-drop tagger~\cite{Butterworth:2008iy}
  conditions, in order
  to enhance the discrimination of the signal over the background
  events.
  %
  For the BDRS mass-drop tagger, we use the {\tt FastJet} default
  parameters of  $\mu = 0.67$ and $y_{\textrm{cut}}= 0.09$.
  %
  Before the mass-drop tagger is applied, the large-$R$ jet
  constituents are reclustered with the Cambridge/Aachen
  algorithm.

@@

  Once we have reconstructed a suitable large-$R$ jet, important
  information can be obtained by looking at the kinematics of
  the small-$R$ subjets.
  %
  These subjets will be obtained by reclustering the constituents
  of the large-$R$ jet, always with the  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter $R=0.3$.
  %
  As we will discuss below, these small-$R$ subjets are the key of
  the $b$-tagging in the boosted category.
%
  \end{itemize}

@@

Given the final state that needs to be reconstructed, the
optimisation of the $b$-tagging capabilities of the
LHC experiments is an essential requirement.
%
The $b$-tagging used in this feasibility study is inspired
in the current and projected ATLAS settings, and below
we also comment on the differences and similarities with the
corresponding CMS strategy.
%
For each type of jet defined above, a different
$b$-tagging strategy is used, as we discuss now.

@@

  These are $b$-tagged as follows.
  %
  If a small-$R$ jet has at least one $b$-quark among their constituents,
  it will be tagged as a $b$-jet with probability $f_b$.
  %
  If no $b$-quarks are found among the constituents
  of this jet, it can be still be tagged as a $b$-jet with
  a mistag rate of $f_l$.
  %
  Note that the probability of tagging a jet is not increased
  if more than one $b$-quark is found among the jet constituents.

@@

The basic idea of this work is the same as in Ref.~\cite{Gouzevitch:2013qca}:
rather than selecting only a specific event topology, such as boosted
or resolved, we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved, which we define now.
%
An important difference as compared to~\cite{Gouzevitch:2013qca}
is that here each category is optimised separately, and then
a exclusive categorisation is defined starting from the topology
where the signal significance is the largest.

@@

  A given event is assigned to the boosted category if it
  contains at least two large-$R$ jets, defined as above,
  and of the tow leading large-$R$ jets have been double-$b$-tagged.
  %
 Therefore, each of these two large-$R$ $b$-tagged jets correspond to the
 decay products of one Higgs boson candidate.

@@

  An event is assigned to the intermediate category if
  at least one large $R$ $b$-tagged jet is present.
  %
  If more than one is found, then the leading large-$R$ jet
  is assigned to be the first Higgs candidate.
  %
  In addition, we require at least two small-$R$ jets,
  which must be separated with respect to the large-$R$
  jet by at least an angular separation of $\Delta R\ge 1.0$.
  %
  The small-$R$ jets must be $b$-tagged as well.
  %
  Then, the second Higgs boson candidate is reconstructed
  by selecting the two small-$R$ jets that minimize the difference
  between the invariant mass of the large-$R$ jet, $m_{lR}$,
  with the invariant mass of the sum of the two small-$R$ jets,
  $m_{sR_1+sR_2}$, that is, ${\rm min}\lp m_{lR} -
  m_{sR_1+sR_2}\rp$.

@@

  Finally, let us define the resolved category.
  %
  An event is assigned to this category
  if at least
  four $b$-tagged small-$R$ jets are present.
  %
  The two Higgs candidates are then reconstructed out of the
  leading four small-$R$ jets in the event
  by considering all possibilities of forming two pairs of jets
  with invariant masses $m_{sR_1+sR_2}$ and 
$m_{sR_3+sR_4}$, respectively, and choosing the configuration that minimizes their difference $|m_{sR_1+sR_2} - m_{sR_3+sR_4}|$. 
  %
  The requirement of restricting this selection to only the four
  leading small-$R$ jets arises from the need to reducing the
  combinatorics from the fakes.
\end{itemize}

In addition, events from each category need to satisfy the usual
Higgs mass window cut, defined by
the following condition
\be
|m_{h,j} - 125| < 40~{\rm GeV} \, , j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of the two Higgs candidates
reconstructed as explained above.
%
We note that this cut is substantially looser as the corresponding
cut used in the ATLAS and CMS $h\to b\bar{b}$ analysis: the motivation
for this is that the MVA will take care of the optimisation of each cut.

In Table~\ref{sec:categorisation} we summarize the definitions of each
category.
%
We emphasize that each category is not exclusive of the others,
and in general there can be a substantial overlap
between the different categories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{c|c|c|c}
    \hline
    &  Boosted  &  Intermediate  &  Resolved  \\
    \hline
    \hline 
    Jet selection  &  $\ge 2$ large-$R$ jets  & 1 large-$R$ jet, $\ge 2$ small-$R$
    jets  &  $\ge$ 4 small-$R$ jets \\
    \hline
    Jet cuts  & $p_T \ge 200$ GeV   &  $p_T \ge 200$ GeV (large-$R$ jet)
    &  $p_T \ge 40$ GeV \\
    &  & $p_T \ge 40$ GeV (small-$R$ jets)   &    \\
    &   $|\eta|<2.0$  & $|\eta|<2.0$ (large-$R$ jet) & $|\eta|<2.5$ \\
    &        & $|\eta|<2.5$ (small-$R$ jet)  &  \\
    \hline
    $b$-tagging  & 2 double-$b$-tags  & 1 double-$b$-tag & 4 single-$b$-tags \\
      &  & and 2 single-$b$-tags  &  \\
    \hline
    \end{tabular}
  \caption{\small Definition of the three event categories used in this
    work, together with the corresponding basic kinematical cuts.
\label{sec:categorisation}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

In section we describe our analysis strategy, and in particular
the classification of each individual event into
different categories according to the event topology.
%
First of all we discuss the settings and nomenclature that
we will use for jet clustering in this work.
%
Then we move to discuss how $b$-tagging is simulated,
following closely the ATLAS techniques.
%
Then we discuss the categorisation of events between different
topologies, and how we can prioritise among them.

@@

In our study we will use different types of jet definitions,
depending of the type of analysis that needs to be performed, optimized
for the various possible event topologies.
%
Therefore, we being by enumerating the choices that we make.
%
The starting point after the parton shower is the clustering
of all final state particles.
%
In this work all the jet reconstruction algorithms
are obtained from the {\tt FastJet} program~\cite{Cacciari:2011ma},
v3.1.0.
%
We adopt the following terminology for the different types of jets
that we will use in this work:

@@

  These are jets  reconstructed with the
  anti-$k_T$ clustering algorithm~\cite{Cacciari:2008gp} with $R=0.4$ radius.
  %
  When reconstructing small-$R$ jets, we will keep only
  those jets with transverse momentum $p_T \ge 40$~GeV
  and pseudo-rapidity $|\eta|<2.5$ within the central
  acceptance of ATLAS and CMS, jets that fail to satisfy these
  conditions are discarded.
  %

@@

  In addition to the basic $p_T$ and $\eta$
  acceptance requirements, large-$R$ jets should also
  satisfy the  BDRS mass-drop tagger~\cite{Butterworth:2008iy}
  conditions, in order
  to enhance the discrimination of the signal over the background
  events.
  %
  For the BDRS mass-drop tagger, we use the {\tt FastJet} default
  parameters of  $\mu = 0.67$ and $y_{\textrm{cut}}= 0.09$.
  %
  As required, before the mass-drop tagger is applied, the large-$R$ jet
  constituents are reclustered with the Cambridge/Aachen
  algorithm.

@@

  Once we have reconstructed a suitable large-$R$ jet, important
  information can be obtained by looking at the kinematics of
  the small-$R$ subjets.
  %
  These subjets will be obtained by reclustering the constituents
  of the large-$R$ jet, always with the  anti-$k_T$ algorithm,
  but this time with a smaller radius parameter $R=0.3$.
  %
  As we will discuss below, these small-$R$ subjets are the key of
  the $b$-tagging in the boosted category.
  %
  Small-$R$ subjets are required to have $p_T > 50$~GeV and $|\eta|<2.5$,
  else they are discarded.

@@

Given the final state that needs to be reconstructed in our case, the
optimisation of the $b$-tagging capabilities of the
LHC experiments is an essential requirement to optimise
the prospects for the observation of Higgs pair production.
%
The $b$-tagging used in this feasibility study is inspired
in the current and projected ATLAS settings, and below
we also comment on the differences and similarities with the
corresponding CMS strategy~\cite{Khachatryan:2011wq,Chatrchyan:2012jua}.
%
For each type of jet defined above, a different
$b$-tagging strategy is used, as we discuss now.

@@

  These are $b$-tagged as follows.
  %
  If a small-$R$ jet has at least one $b$-quark among their constituents,
  it will be tagged as a $b$-jet with probability $f_b$.
  %
  If no $b$-quarks are found among the constituents
  of this jet, it can be still be tagged as a $b$-jet with
  a mistag rate of $f_l$.
  %
  Note that the probability of tagging a jet is not modified
  if more than one $b$-quark is found among the jet constituents.

  We only attempt to $b$-tag the four (two) hardest small-$R$ jets
  in the resolved (intermediate) category.
  %
  We find that
  trying to $b$-tag all
  small-$R$ jets that satisfy the acceptance cuts worsens the performance
  due to combinatorics.

@@

The basic idea of this work is similar to that in Ref.~\cite{Gouzevitch:2013qca}:
rather than concentrating only a specific event topology, such as boosted
or resolved, we aim to consistently combine the information from
the three possible event topologies: boosted, intermediate and
resolved, which we define now.
%
An important difference as compared to~\cite{Gouzevitch:2013qca}
is that here each category is optimised separately, and then
a exclusive categorisation is defined starting from the topology
where the signal significance is the best.

@@

  A given event is assigned to the {\it boosted} category if it
  contains at least two large-$R$ jets, defined as above,
  and where  the two leading large-$R$ jets have been
  successfully double-$b$-tagged.
  %
 Therefore, each of these two large-$R$ double-$b$-tagged jets corresponds to the
 decay products of one Higgs boson candidate.

@@

  An event is assigned to the {\it intermediate} category if
  at least one large $R$ double-$b$-tagged jet is present.
  %
  If more than one is found, then the leading large-$R$ jet
  is assigned to be the first Higgs candidate.
  %
  In addition, we require at least two small-$R$ jets,
  which must be separated with respect to the large-$R$
  jet by at least an angular separation of $\Delta R\ge 1.0$.
  %
  These small-$R$ jets must be $b$-tagged as well.

  Then, the second Higgs boson candidate is reconstructed
  by selecting the two $b$-tagged small-$R$ jets that minimize the difference
  between the invariant mass of the large-$R$ jet, $m_{lR_1}$,
  with the invariant mass of the sum of the two small-$R$ jets,
  $m_{sR_1+sR_2}$, that is, ${\rm min}\lp m_{lR} -
  m_{sR_1+sR_2}\rp$, where $lR_1$ indicates the leading large-$R$ jet
  and $sR_1$ and $sR_2$ indicate the small-$R$ jets.m

@@

  Finally, let us define the {\it resolved} category.
  %
  An event is assigned to this category
  if at least
  four $b$-tagged small-$R$ jets are present.
  %
  The two Higgs candidates are then reconstructed out of the
  leading four small-$R$ jets in the event
  by considering all possibilities of forming two pairs of jets
  with invariant masses $m_{sR_1+sR_2}$ and 
$m_{sR_3+sR_4}$, respectively, and choosing the configuration that minimizes their difference $|m_{sR_1+sR_2} - m_{sR_3+sR_4}|$. 
  %
  The requirement of restricting this selection to only the four
  leading small-$R$ jets arises from the need to reducing the
  combinatorics from the fakes.
\end{itemize}
Note that in principle these three categories are not exclusive:
a given even can be assigned to more than one category, for
example, satisfying both the requirements of the boosted and resolved
categories at the same time.

Once a pair of Higgs boson candidates has been identified,
then events from each category are required to satisfy the usual
Higgs mass window cut, defined by
the following condition
\be
|m_{h,j} - 125| < 40~{\rm GeV} \, , j=1,2 \, ,
\ee
where $m_{h,j}$ is the invariant mass of the two Higgs candidates
reconstructed as explained above.
%
We note that this cut is substantially looser as the corresponding
cut used in the typical ATLAS and CMS $h\to b\bar{b}$
analyses~\cite{Aad:2012gxa,Chatrchyan:2013zna}: the motivation
for this is that the MVA will take care of the optimisation of these
and other cuts that are not explicitly performed.

In Table~\ref{sec:categorisation} we summarize the definitions of each
category.
%
We emphasize again that each category is not exclusive of the others,
and in general there can be a substantial overlap
between the different categories.
%
In particular, this overlap is a direct consequence that
the different categories are optimised separately, as opposed
to the approach of~\cite{Gouzevitch:2013qca}, where optimisation
was performed simultaneously in the combination of exclusive categories.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
  \centering
  \small
  \begin{tabular}{c|c|c|c}
    \hline
    &  Boosted  &  Intermediate  &  Resolved  \\
    \hline
    \hline 
    Jet selection  &  $\ge 2$ large-$R$ jets  & 1 large-$R$ jet, $\ge 2$ small-$R$
    jets  &  $\ge$ 4 small-$R$ jets \\
    \hline
    Jet cuts  & $p_T \ge 200$ GeV   &  $p_T \ge 200$ GeV (large-$R$ jet)
    &  $p_T \ge 40$ GeV \\
    &  & $p_T \ge 40$ GeV (small-$R$ jets)   &    \\
    &   $|\eta|<2.0$  & $|\eta|<2.0$ (large-$R$ jet) & $|\eta|<2.5$ \\
    &        & $|\eta|<2.5$ (small-$R$ jet)  &  \\
    \hline
    $b$-tagging  & 2 double-$b$-tags  & 1 double-$b$-tag & 4 single-$b$-tags \\
    &  & and 2 single-$b$-tags  &  \\
       & $\ge 2$ small-$R$ subjets  & &  \\
    &  $p_T^{\rm sj}\ge 50$ GeV &   &  \\
    \hline
    \end{tabular}
  \caption{\small Definition of the three event categories used in this
    work, together with the corresponding basic kinematical cuts.
    %
    The categories are not exclusive, and each event can satisfy the
    requirements of more than one category at the same time.  
\label{sec:categorisation}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

As we will show now, the selection requirement and the kinematic cuts
for each of the categories in Table~\ref{sec:categorisation} have
been optimised independently.
%
First of all we have performed a inclusive analysis, and optimised
$S/\sqrt{B}$ in each of the three categories separately, including
the MVA.
%
We find that, in terms of signal significance, the ordering of
the categories is
\begin{enumerate}
\item Boosted 
\item Intermediate
  \item Resolved
  \end{enumerate}
and therefore for the final analysis we turn to an exclusive approach:
if an event satisfies the boosted requirements, it is kept for
these category, else we check if it suits the intermediate
requirements, and failing also this, the event is attempted
in the resolved categories.
%
The resulting exclusive event samples are then separately processes
through the MVA, as discussed in the next section.
%
This allows us to combine the significance of the dominant topology
(boosted) with that of the sub-leading categories, intermediate
and resolved.
